---
title: "Forecasting Section 6.7"
author: "Seema Rani Kanuri"
date: "May 3, 2017"
output: html_document
---


#Question 1:
Show that a 3×53×5 MA is equivalent to a 7-term weighted moving average with weights of 0.067, 0.133, 0.200, 0.200, 0.200, 0.133, and 0.067.

```{r}
library(forecast)
library(fpp)
library(TTR)
```


```{r}
values=c(0.067, 0.133, 0.200, 0.200, 0.200, 0.133,0.067)
moveavg = ma(values, order = 5)
weightedmoveavg=WMA(values,n = 5)
sum(na.omit(moveavg))
sum(na.omit(weightedmoveavg))
```


#Question 2:
The data below represent the monthly sales (in thousands) of product A for a plastics manufacturer for years 1 through 5 (data set plastics).

###a. Plot the time series of sales of product A. Can you identify seasonal fluctuations and/or a trend?

```{r}
library(fma)
library(fpp)
```

```{r}
ts <- plastics
plot(plastics)
```
There are seasonal fluctuation with sales peaking in summer and reaching troughs in winter. There is a positive trend throughout the series.


###b. Use a classical multiplicative decomposition to calculate the trend-cycle and seasonal indices.
```{r}
fit <- decompose(ts, type="multiplicative")

trend_indices <- fit$trend
seasonal_indices <- fit$seasonal
```


###c. Do the results support the graphical interpretation from part (a)?
```{r}
```
The graph indicates that the summer months should have higher seasonal indices than the winter months and this is indeed the case.


###d. Compute and plot the seasonally adjusted data.
```{r}
first <- seasadj(fit)
plot(first)
```


##e. Change one observation to be an outlier (e.g., add 500 to one observation), and recompute the seasonally adjusted data. What is the effect of the outlier?
```{r}
ts2 <- ts
ts2[13] = ts2[13] + 500
fit2 <- decompose(ts2, type="multiplicative")
second <- seasadj(fit2)
plot(second)
```
The outlier causes a spike in the graph at the month it was inserted.


###f. Does it make any difference if the outlier is near the end rather than in the middle of the time series?
```{r}
ts3 <- ts
ts3[29] = ts3[29] + 500

fit3 <- decompose(ts3, type='multiplicative')

third <- seasadj(fit3)

ts4 <- ts
ts4[51] = ts4[51] + 500

fit4 <- decompose(ts4, type='multiplicative')

fourth <- seasadj(fit4)

plot(third)
plot(fourth)
```
It appears that it doesn't matter where the outlier occurs. Whereever it is inserted we see a similar (perhaps identical) sized spike.

###g. Use a random walk with drift to produce forecasts of the seasonally adjusted data.
```{r}
fit5 <- rwf(first, drift=TRUE)
```

###h. Reseasonalize the results to give forecasts on the original scale.
```{r}
fit5 <- stl(ts, t.window=15, s.window="periodic", robust=TRUE)
fit6 <- forecast(fit5, method="naive")
```



#Question 3:
Figure 6.13 shows the result of decomposing the number of persons in the civilian labor force in Australia each month from February 1978 to August 1995.

###a. Write about 3-5 sentences describing the results of the seasonal adjustment. Pay particular attention to the scales of the graphs in making your interpretation.
```{r}

```
Isolating the trend component from the seasonal component shows that the trend has increased throught the majority of the time frame, with a few stationary bperiods occuring in the early 90s. The monthly breakdown of the seasonal component shows that a few months show greater velocities in their variations than other months. 

###b. Is the recession of 1991/1992 visible in the estimated components?
```{r}

```
Yes the large residuals during that time period capture this information.
