---
title: "Forecasting Section 8.11"
author: "Seema Rani Kanuri"
date: "May 3, 2017"
output: html_document
---


#Question 5:
Use R to simulate and plot some data from simple ARIMA models.

###a. Use the following R code to generate data from an AR(1) model with ??1=0.6??1=0.6 and ??2=1??2=1. The process starts with y0=0y0=0.
```{r}
library(fpp)
library(knitr)
```

```{r}
Alpha <- ts(numeric(100))
Beta<- rnorm(100)
for(i in 2:100)
  Alpha[i] <- 0.6*Alpha[i-1] + Beta[i]
```




###b. Produce a time plot for the series. How does the plot change as you change ??1??1?
```{r}
plot(Alpha, main="TIMESERIES GRAPH", xlab="time")
Gamma <- function(n.obs, phi, seed.nr){
  set.seed(seed.nr)
  Alpha <- ts(numeric(n.obs))
  Beta <- rnorm(n.obs)
  for (i in 2:n.obs)
    Alpha[i] <- phi*Alpha[i-1] + Beta[i]
  return(Alpha)
}
```

```{r}
plot(Gamma(100, 0.9, 2), main= "time series", xlab="times")
lines(Gamma(100, 0.9, 2), col="green")
```



###c. Write your own code to generate data from an MA(1) model with ??1=0.6??1=0.6 and ??2=1??2=1.
```{r}
Charlie <- function(n.obs, theta, seed.nr){
  set.seed(seed.nr)
  Alpha <- ts(numeric(n.obs))
  Beta <- rnorm(n.obs)
  for (i in 2:n.obs)
    Alpha[i] <- theta*Alpha[i-1] + Beta[i]
  return(Alpha)
}
```



##d. Produce a time plot for the series. How does the plot change as you change ??1??1?
```{r}
plot(Charlie(100, 0.6, 2), main="time series", xlab="time")
```


###e. Generate data from an ARMA(1,1) model with ??1??1 = 0.6 and ??1=0.6??1=0.6 and ??2=1??2=1.
```{r}
Xulu <- function(n.obs, theta, seed.nr){
  set.seed(seed.nr)
  Alpha <- ts(numeric(n.obs))
  Beta <- rnorm(n.obs)
  for (i in 2:n.obs)
    Alpha[i] <- phi*Alpha[i-1] + theta*Beta[i-1] + Beta[i]
  return(Alpha)
}
```



###f. Generate data from an AR(2) model with ??1=???0.8??1=???0.8 and ??2=0.3??2=0.3 and ??2=1??2=1. (Note that these parameters will give a non-stationary series.)
```{r}
Echo <- function(n.obs, phi1, phi2, seed.nr){
  set.seed(seed.nr)
  Alpha <- ts(numeric(n.obs))
  Beta <- rnorm(n.obs)
  for(i in 3: n.obs)
    Alpha[i] <- phi1*Alpha[i-1] + phi2*Alpha[i-2] + Beta[i]
  return(Alpha)
}
```

```{r}
plot(Echo(100, -0.9, 0.3, 2), main="timeseries", xlab="time")
```



###g. Graph the latter two series and compare them.

```{r}
par(mfrow=c(2,1))
plot(Charlie(100, 0.6, 2), main="time series", xlab="time")
plot(Echo(100, -0.9, 0.3, 2), main="timeseries", xlab="time")
dev.off()
```



#Question 6;
Consider the number of women murdered each year (per 100,000 standard population) in the United States (data set wmurders).

###a. By studying appropriate graphs of the series in R, find an appropriate ARIMA(p,d,qp,d,q) model for these data.
```{r}
data(wmurders)
tsdisplay(wmurders)
Alpha1 <- diff(wmurders)
tsdisplay(Alpha1)
adf.test(Alpha1)
kpss.test(Alpha1)
```

```{r}
wmurders.d2 <- diff(diff(wmurders))

tsdisplay(wmurders.d2)
adf.test(wmurders.d2)
kpss.test(wmurders.d2)
```

```{r}
test.arima <- function(t.series, order){
  Alpha3 <- data.frame(model=paste0("ARIMA(",paste0(order, collapse=","), ")"),AICc=Arima(t.series, order=order)$aicc)
  return(Alpha3)
}
Alpha2 <- expand.grid(c(0, 1, 2),c(1, 2),c(0, 1, 2))
Alpha3.list <- apply(Alpha2, MARGIN=1, FUN=function(x) {test.arima(wmurders, x)})
Alpha3 <- do.call(rbind, Alpha3.list)
kable(Alpha3)
```

```{r}
Alpha4 <- Arima(wmurders, order=c(0, 1, 2))
tsdisplay(residuals(Alpha4), lag.max=20)
Box.test(residuals(Alpha4), lag=24, fitdf=4, type="Ljung")
```




###b. Should you include a constant in the model? Explain.
```{r}

```
No. A constant introduces drift into the model, which we do not appear to have in these data.




###c. Write this model in terms of the backshift operator.
```{r}

```
$$ (1 - B)^2 y_t = (1 + \theta_1B + \theta_2B^2)e_t$$



###d. Fit the model using R and examine the residuals. Is the model satisfactory?
```{r}
Alpha4 <- Arima(wmurders, order=c(0, 1, 2))
tsdisplay(residuals(Alpha4), lag.max=20)
```




###e. Forecast three times ahead. Check your forecasts by hand to make sure you know how they have been calculated.
```{r}
Forecast <- forecast(Alpha4, h=3)
Forecast$mean
toforecast <- 3

Alpha5 <- Alpha4$x
Alpha6 <- Alpha4$residuals
theta1 <- as.numeric(Alpha4$coef['ma2'])
theta2 <- as.numeric(Alpha4$coef['ma1'])
```

```{r}
for (h in 1:toforecast){
  n <- length(Alpha5)
  Alpha7 <- 2 * Alpha5[n] - Alpha5[n - 1] + theta1 * Alpha6[n] + theta2 * Alpha6[n - 1]
  Alpha5 <- c(Alpha5, Alpha7)
  Alpha6 <- c(Alpha6, 0)
}

```

```{r}
f <- Alpha5[(length(Alpha5) - toforecast + 1):length(Alpha5)]

plot(Forecast)
lines(Alpha4$x - Alpha4$residuals, col='blue')
points(c(2005, 2006, 2007), f, col='red')
```



###f. Create a plot of the series with forecasts and prediction intervals for the next three periods shown.
```{r}
plot(Forecast)
```



##sss#g. Does auto.arima give the same model you have chosen? If not, which model do you think is better?
```{r}
auto.arima(wmurders)
```
